Estimating SARS-CoV-2 transmission in England from randomly sampled prevalence surveys

### 2023-11-13 SEIR.tex
Various edits
Fix formatting

\begin{document}
\ifSubfilesClassLoaded{
    \setcounter{chapter}{7}
}
\chapter{Mechanistic modelling of incidence and transmission} \label{SEIR}

@@

The basic SIR model implies that the distribution of the time spent in the infected compartment is exponential~\autocite[96]{keelingModeling}.
This is an unrealistic assumption and leads to underestimation of the reproduction number~\autocites{lloydRealistic}{wearingAppropriate}.
A standard extension to the SIR model is to allow the infectious period to follow a gamma distribution.
This is done by adding multiple infected compartments (see \cref{SEIR:fig:SIR-gamma}).
The system can now be written as follows~\autocite[94]{keelingModeling}:
\begin{align}
\frac{ds}{dt} &= -\beta si \\
\frac{di_1}{dt} &= \beta si - n\gamma i_1 \\
\frac{di_2}{dt} &= n\gamma i_1 - n \gamma i_2 \\
&\vdots \nonumber \\
\frac{di_n}{dt} &= n\gamma i_{n-1} - n \gamma i_n \\
\frac{dr}{dt} &= n\gamma i_n
\end{align}
where $n$ is the number of infected compartments.
The infectious period is now distributed $\GamDist(n, n\gamma)$.
A gamma distribution with the first parameter equal to 1 is an exponential distribution, connecting this model to the basic SIR model in \cref{SEIR:sec:SIR}.
The distribution for the time spent in the infected compartments tends to a constant (\ie the same for all individuals) as $n \to \infty$~\autocite{lloydRealistic}.
The form of these equations is chosen so that $1/\gamma$ remains the mean infectious period.

Using a two parameter distribution allows the first two moments of the infectious period to be specified.
This is normally sufficient to give realistic dynamics within a mean-field model.

@@

@@

\subsection{MCMC}

>>>>

\begin{document}
\ifSubfilesClassLoaded{
    \setcounter{chapter}{7}
}

\chapter{Mechanistic modelling of incidence and transmission} \label{SEIR}

@@

The basic SIR model implies that the distribution of the time spent in the infected compartment is exponential~\autocite[96]{keelingModeling}.
This is an unrealistic assumption and leads to underestimation of the reproduction number~\autocites{lloydRealistic}{wearingAppropriate}.
A standard extension to the SIR model is to allow the infectious period to follow a gamma distribution.
This is done by adding multiple infected compartments (see \cref{SEIR:fig:SIR-gamma}).
The system can now be written as follows~\autocite[94]{keelingModeling}:
\begin{align}
\frac{ds}{dt} &= -\beta si \\
\frac{di_1}{dt} &= \beta si - n\gamma i_1 \\
\frac{di_2}{dt} &= n\gamma i_1 - n \gamma i_2 \\
&\vdots \nonumber \\
\frac{di_n}{dt} &= n\gamma i_{n-1} - n \gamma i_n \\
\frac{dr}{dt} &= n\gamma i_n
\end{align}
where $n$ is the number of infected compartments.
The infectious period is now distributed $\GamDist(n, n\gamma)$.
Gamma distributions are a realistic model for a variety of diseases~\autocite{wearingAppropriate}.
A gamma distribution with the first parameter equal to 1 is an exponential distribution, connecting this model to the basic SIR model in \cref{SEIR:sec:SIR}.
The distribution for the time spent in the infected compartments tends to a constant (\ie the same for all individuals) as $n \to \infty$~\autocite{lloydRealistic}.
The form of these equations is chosen so that $1/\gamma$ remains the mean infectious period.

@@

I use previously developed estimates of how contact matrices change over time.
These matrices were produced for \textcite{birrellRealtime}.
These start with estimates of contact in each setting, based on \textcite{vanleeuwenAugmenting}.
These use data from the POLYMOD study~\autocite{mossongSocial} and UK Time Use Survey (UKTUS)~\autocite{UKTUS}.
POLYMOD asked a representative sample of the UK population about the contacts they make each day, including information on age and location (home, work, school, leisure, transport, or other) of these contacts.
UKTUS asked a representative sample of the UK population about how much time they spent doing various activities each day.
\Textcite{vanleeuwenAugmenting} combined these two datasets to estimate the rate of contacts between age groups during each activity.
\Textcite{birrellRealtime} then modify these contact rates based on measured changes in time use during the pandemic.
These changes are measured using Google mobility data~\autocite{googleCOVID19} and school attendance data~(unpublished).
Google mobility data measures the location of Android users over time, relative to a pre-pandemic baseline.
These data sources are then used to down or upweight the number of contacts in each setting, as estimated by \textcite{vanleeuwenAugmenting}.
For mathematical details see \textcite{vanleeuwenAugmenting} and \textcite[supplementary material]{birrellRealtime}.

@@

\subsection{MCMC} \label{SEIR:sec:MCMC}
